---
title: "A function to round counts to nearest 5 in gtsummary baseline tables"
author: "Zheer Kejlberg Al-Mashhadi"
date: "2023-11-04"
output: github_document
---

### Definition of the function and use example

<br>
  
#### 1) Load dependencies

```{r, results='hide', error=FALSE, warning=FALSE, message=FALSE}
library(gtsummary) # For creating a baseline characteristics table
library(tidyverse) # For data wrangling and misc.
```
  
<br>
  
<br>
  
#### 2) Define the function

```{r}

round_5_gtsummary <- function(table) {
  
  round_5 <- function(x) { round(x/5)*5 }
  
  round_5_get_summary <- function(x, N, decimals = 1) {
    x <- str_remove(x, " \\([<]*[0-9]*[,]*[0-9]*[.]*[0-9]*%\\)$")
    x <- as.numeric(str_remove(x, ","))
    
    if (x > N-5) {
      N <- round_5(N)
      return(paste0(">", N-5, "(>", round((N-5)/N*100, decimals), "%)"))
    } else if (x >= 5) {
      return(paste0(round_5(x), " (", round(round_5(x)/round_5(N)*100,decimals),"%)"))
    } else {
      return(paste0("<", 5," (<", round(5/round_5(N)*100,decimals),"%)"))
    }
  }
  
  body <- table$table_body
  stats_column_indices <- which(grepl("^stat_", colnames(body)))
  
  Ns <- table$table_styling$header$modify_stat_n[c(stats_column_indices)]
  table$table_styling$header$label[c(stats_column_indices)] <- paste0("**", table$table_styling$header$modify_stat_level[c(stats_column_indices)], "**", ", N = ", round_5(Ns))
  
  for (column_no in stats_column_indices) {
    column <- pull(body, column_no)
    cat_indices <- (body$var_type == "categorical" | body$var_type == "dichotomous" | body$label == "Unknown") & !is.na(body$stat_1)
    N <- table$table_styling$header$modify_stat_n[column_no]
    column[cat_indices] <- sapply(column[cat_indices], round_5_get_summary, N = N)
    table$table_body[column_no] <- column
  }
  return(table)
}
```
  
<br>

<br>
  
#### 3) Run it on a tbl_summary (unweighted) or tbl_svysummary (weighted) table


For comparison, here's a table without rounding
```{r}

trial %>% 
  tbl_summary(by = grade, include = c(trt, age, stage))

```
  
<br>
  
Now, the same table but with all counts rounded to nearest 5 (and all proportions adjusted accordingly)
```{r}

trial %>% 
  tbl_summary(by = grade, include = c(trt, age, stage)) %>%
  round_5_gtsummary()

```
  
<br>
  
Finally, the function can also be applied to weighted data (here shown with and without rounding)
```{r, error=FALSE, warning=FALSE, message=FALSE}

library(WeightIt) # for calculating weights
weighted_trial <- trial %>% 
  mutate(w = weightit(grade ~ trt + age + stage, estimand = "ATT", focal = "I")$weights) %>%
  survey::svydesign(~1, data = ., weights = ~w)

weighted_trial %>%
  tbl_svysummary(by = grade, include = c(trt, age, stage))

weighted_trial %>%
  tbl_svysummary(by = grade, include = c(trt, age, stage)) %>% 
  round_5_gtsummary()

```
